### 2018-06-12 MSc_Marko_Docevski_tekst.tex
Updates from ShareLaTeX

\cite{Johnson2017} Product of Experts
Hinton (2002) proposed a system known as Product of Ex- perts (PoE) for combining multiple models of the same data. To combine a series of N experts with parame- ters Θ1,Θ2, . . . ,ΘN and associated probability distribu-
tions f1(·|Θ1), f2(·|Θ2), . . . , fN(·|ΘN), one can renormal- ize the product of their individual distributions.
Notice the similarity of this expres- sion to the definition of conditional probability: p(A|B) = p(A∩B) p(B) . In fact, this product can be interpreted as the con-
ditional probability that all experts choose x, given that all experts choose the same vector from the data space.
We use a pair of encodings, motivated
by the observation that good jazz melodies have both inter- esting contours as well as pitches that are sonorous with the chord progression. One encoding is based on intervals be- tween adjacent notes and the other is based on harmonic relationships with the current chord in the chord progres- sion. Each encoding is processed by a separate network component, and each component produces a candidate note probability distribution. These distributions are then com- bined using Product-of-Experts (Hinton, 2002) to produce a final distribution over notes that can be trained using the maximum-likelihood criterion.
The network is de- signed to receive the previous note played as input at each step, producing as output a set of probabilities for which note to play at the next time step. The network is trained using the maximum-likelihood objective (i.e. choose the network model that assigns the highest probability to the true training data).
In order to fully develop a LSTM-based model for jazz melodies, we needed to choose a representation for the musical data for the network to understand. Furthermore, dif- ferent representations might lead to different types of be- havior. For instance, if we simply represented the notes by a bit-vector where each bit corresponded to a single note, the model would not be able to generalize well to different in- tervals and other patterns, as it would have to learn relation- ships between each note separately. Franklin (2004) found that representations that were motivated by musical relation- ships led to better performance than more naive approaches.
To accomplish these goals, we decided to use two encod-
ings, one that focuses on intervals between successive notes, and the other that focuses on intervals of notes relative to the chord progression. We thus split our network into two network modules. The interval expert module receives an interval-based encoding, while the chord expert receives the chord-based encoding. This enables the network to learn particular relationships relative to the current trajectory of the melody (as learned by the interval expert) and also rela- tive to the current chord progression (as learned by the chord expert). Each of these modules produces a probability dis- tribution over which note to choose at the next time step. This probability distribution represents the level of belief that those notes should be chosen to play next. The encod- ings used are depicted in F2
Each expert receives as input:
1. the note chosen at the previous time step, encoded using the corresponding expert’s output format (so the interval expert receives the previous interval jump, and the chord expert receives the previous relative pitch class).
2. a beat vector giving the position of the current time step in a measure.
3. a position vector giving the position of the previous note relative to the upper and lower bounds.
4. a chord vector giving the notes of the current chord rela- tive to the expert’s current position.
Training
Our network is trained using the cross-entropy between the network predictions at each time step and the correct notes chosen in the training data. Equivalently, we want to maxi- mize (the log-likelihood of) the probability that the network outputs the training data perfectly, since we want the net- work to output improvisations that are similar to the training data. We can express the probability of a whole melodic seg- ment as a product of conditional distributions at each time step, i.e. the probability of generating a segment x from pa- rameters Θ can be factored as To train our network, we can then compute the gradient of the loss with respect to our parameters Θ by performing backpropagation through time (Werbos, 1990). In our experiments, each expert subnetwork was imple-
mented as two LSTM layers, each with 300 nodes. Dur- ing training, dropout of 0.5 was applied to the non-recurrent connections between layers (Srivastava et al., 2014; Pham et al., 2014). We used the ADAM optimizer, introduced by Kingma and Ba (2014), to automatically set the learning rate for our training procedure.
Generation
After our model is fully trained, we can use it to create new melodies from the learned probability distribution. At each time step, starting from the beginning of the piece, we com- pute the probability distribution output by our model. We then choose the next note to play proportionally to the prob- ability assigned to that note by our model. This note is then fed back into the model as the input for the next time step, and a probability distribution for the next time step is computed. This process repeats until we have created a full melody segment. Optionally, we can modify the probability distribution be-
fore sampling from it, in order to modify or constrain the output of the network. The first way to modify the distribu- tion is to vary the ratios between note probabilities.
Small values of x make the model more “adventurous” by causing the probabilities to become more similar to each other, and large values of x will cause the model to be “con- servative” by sampling high-probability notes more often and low-probability notes less often. Setting x = 1 results in the original probability distribution.

@@

\cite{Mehri2016} Hence, our contribution is threefold:
1. We present a novel method that utilizes RNNs at different scales to model longer term de- pendencies in audio waveforms while training on short sequences which results in memory efficiency during training.
2. We extensively explore and compare variants of models achieving the above effect.
3. We study and empirically evaluate the impact of different components of our model on three audio datasets. Human evaluation also has been conducted to test these generative models.
SampleRNN helps to address this challenge by using a hierarchy of modules, each operating at a different temporal resolution. The lowest module processes individual samples, and each higher module operates on an increasingly longer timescale and a lower temporal resolution. Each module conditions the module below it, with the lowest module outputting sample-level predictions. The entire hierarchy is trained jointly end-to-end by backpropagation.
Rather than operating on individual samples, the higher-level modules in SampleRNN operate on non-overlapping frames of FS(k) (“Frame Size”) samples at the kth level up in the hierarchy at a time (frames denoted by f(k)). Each frame-level module is a deep RNN which summarizes the history of its inputs into a conditioning vector for the next module downward.
Because different modules operate at different temporal resolutions, we need to upsample each
vector c at the output of a module into a series of r(k) vectors (where r(k) is the ratio between the temporal resolutions of the modules) before feeding it into the input of the next module downward (Eq. 6). We do this with a set of r(k) separate linear projections.
Because different modules operate at different temporal resolutions, we need to upsample each
vector c at the output of a module into a series of r(k) vectors (where r(k) is the ratio between the temporal resolutions of the modules) before feeding it into the input of the next module downward (Eq. 6). We do this with a set of r(k) separate linear projections.
The lowest module (tier k = 1; Eqs. 7–9) in the SampleRNN hierarchy outputs a distribution over a sample xi+1, conditioned on the FS(1) preceding samples as well as a vector c(k=2)
from the i
next higher module which encodes information about the sequence prior to that frame. As FS(1) is usually a small value and correlations in nearby samples are easy to model by a simple memoryless module, we implement it with a multilayer perceptron (MLP) rather thanRNNwhich slightly speeds up the training. 
We use a Softmax because we found that better results were obtained by discretizing the audio signals (also see van den Oord et al. (2016)) and outputting a Multinoulli distribution rather than using a Gaussian or Gaussian mixture to represent the conditional density of the original real-valued signal. When processing an audio sequence, the MLP is convolved over the sequence, processing We use a Softmax because we found that better results were obtained by discretizing the audio signals (also see van den Oord et al. (2016)) and outputting a Multinoulli distribution rather than using a Gaussian or Gaussian mixture to represent the conditional density of the original real-valued signal. When processing an audio sequence, the MLP is convolved over the sequence, processing
We enable efficient training of our recurrent model using truncated backpropagation through time, splitting each sequence into short subsequences and propagating gradients only to the beginning of each subsequence. We experiment with different subsequence lengths and demonstrate that we are able to train our networks, which model very long-term dependencies, despite backpropagating through relatively short subsequences.
We implemented the WaveNet architecture as described in Oord et al. (2016). Ideally, we would have liked to replicate their model exactly but owing to missing details of architecture and hyper- parameters, as well as limited compute power at our disposal, we made our own design choices so that the model would fit on a single GPU while having a receptive field of around 250 milliseconds
Apart from reporting NLL, we conducted AB preference tests for random samples from four models trained on the Blizzard dataset. For unconditional generation of speech which at best sounds like mumbling, this type of test is the one which is more suited. Competing models were the RNN, SampleRNN (2-tier), SampleRNN (3-tier), and our implementation of WaveNet. The rest of the models were excluded as the quality of samples were definitely lower and also to keep the number of pair comparison tests manageable.

@@

\cite{Tikhonov2017} A crucial feature of LSTM network that makes it extremely attractive in this context is that LSTM shows significantly better results when dealing with time lags of unknown size between important events [5]. This comparable insen- sitivity to gap length gives a unique advantage to LSTMs over hidden Markov models, alternative recurrent neural networks and other sequence learning meth- ods when algorithm works withmusic. Music patterns can be temporally complex and LSTMs seem to be apt to capture this complexity to a high extent. highway network cell?! A crucial feature of LSTM network that makes it extremely attractive in this context is that LSTM shows significantly better results when dealing with time lags of unknown size between important events [5]. This comparable insen- sitivity to gap length gives a unique advantage to LSTMs over hidden Markov models, alternative recurrent neural networks and other sequence learning meth- ods when algorithm works withmusic. Music patterns can be temporally complex and LSTMs seem to be apt to capture this complexity to a high extent. They focus on monotonic music. Monotonic melody could be represented as a sequence of characters which allows us to apply a number of approaches and algorithms that proved to be successful for automated text generation. This significantly reduces dimenstionality (thnik of a dataset of all characters vs a dataset of all valid words). A char-based approach in which an algorithm generates texts letter by letter is closer to the case of monotone music generation. 4 GB dataset of MIDI FILES!!! JEZUZ Dataset preprocessing: 1. Spliting the midi tracks and filtering only the desired ones by heuristic. 2. Removed strenght/velocity of the note being played, to focus on the melodic pattern determined by the pitches and temporal parameters of the notes and pauses in between and not performance nuances. 3. In order to make the learning state space denser we have centered the pitches throughout the dataset transposing median pitch of every track to the 4th octave. 4. We also normalized the pauses throughout the dataset in the following manner. For each track we have calculated a median pause. It is only to be expected that absolute majority of the pauses in the track were equal to the median pause multiplied with a rational coefficient (naturally 1/2 and 3/2 were especially popular in the majority of the tracks). We counted all possible pauses in every track and left only the tracks that had 11 different values of the pauses or less (the median + most popular pauses on each side of it). The tracks with higher variety of pauses were not included in the final dataset.
5. Finally to make the input diverse enough we have filtered the tracks with exceedingly small entropy. Result was 15+ thousand normalized tracks. For each note we were building a note embedding that corresponded to the pitch of the note, an octave embedding that corresponded to the octave of the note and a delay embedding that corresponded to the length of the note. We were using this three embeddings and meta-information of a given MIDI track to build a concatenated note representation that was used as an input for training throughout this paper. basic principles that could be applied to a variety of LSTMs developed for language modeling: – The input words are encoded by 1-of-K coding where K is the number of words in the vocabulary, – At the output layer, a softmax activation function is used to produce cor- rectly normalized probability values, – The cross entropy error which is equivalent to maximum likelihood is used as a training criterion. LSTM language model in the context of music generation. A weakness of LM is that it does not capture global features in an interpretable way. Models like VAE or Reccurent Highway Network Pay more attention to the macro structure of the track. Recurrent Highway Networks extend the LSTM architecture to allow step-to-step transition depths larger than one. Reorganized LSTM language model that is described in Figure 2 and proposed the architecture shown in Figure 3 that is a version of VAE called variational recurrent autoencoder. Here the first network (encoder) compresses the given track into a latent vector that works as a bottleneck. The second network (decoder) learns to reconstruct the melody out of a latent repre- sentation. This approach stimulates the network to work with a macrostructure of the track due to the low dimensionality of the latent vector. Naturally, there is a trade-off between the potential of the network to capture the macro-structure and its possibility to generate locally diverse melodies. One would like to propose an architecture that could combine both these features and would balance local diversity with global structure. Variational autoencoder scheme for music generation. VRASH VRASH in a sense combines a language model and variational recurrent autoencoder in order to increase the performance on the data with varying input length. VRASH architecture is principally described in Figure 4. Here analo- gously to the scheme on Figure 3 the decoder tries to reconstruct the track out of the latent vector, but this vector is distorted with a variational bayesian noise. The decoder also uses the previous outputs as additional inputs. It ”listens” to the notes that it has composed already and uses them as additional ”historic” inputs. Variational Recurrent Autoencoder Supported by History (VRASH) scheme for music generation. There is no objective comparison metric for art generation algorithms. Most aprochest tend to use peer-review systems, and the number of humans significantly varies depending on the research, from a couple dozen to hundreds. Generally speaking, with an ever growing interest of computer scien- tists to art-generating algorithms one would expect the development of some rigorous art metrics to become a specific task within the interdisciplinary focus of arts and sciences. They chose to compare the proposed architectures with respect to the cross-entropy that is commonly used as a loss-term in such task and share their personal suubjective opinion on the outputs. Despite not having a much better cross-entropy value than a Language Model (LSTM based model) the authors claim that a autoencoder based model captures macro structure better, at least in their subjective opinion. All 3 of their experiment architecture achived interesting results. The first problem that cocurs in many generative models is the tendency to repeat a certain note. It turned out to be more prominent in the LM, while VAE and VRASH delt with it a little better. Secondly descpite being designed to capture the macrostructure of the track, they do not always provide one like in a human writen track.  The authors still believe that VRASH is the way to go. It could be. They suggest improvement in: 1. A better structure classification that coud be used within a meta-information input for every track 2. Develop other architectures that would be capable of captureing repetitive melodic structures that are placed on varying distances within a given track. Model should be relatively easy to implement and train.

@@

@@

>>>>

@@

Во системот наречен „Производ на експерти“ предложен во \cite{Johnson2017} се користат два прости јазични модели во тандем, тренирани на истото податочно множество од џез песни, со различен перпектива врз истото. Двата модела се едноставни рекурентни длабоки невронски мрежи, од кои едниот ги гледа песните како секвенца од релативно движење на мелодијата, т.е. при секој чекор ја гледа абсолутната разлика во тонот меѓу последователни ноти, додека другиот ја следи хармониската улога на нотата во тековниот акорд од прогесијата на акорди. Двата модели при секој чекор моделираат веројатностна дистрибуција и може едноставно да се тренираат со оптимизациски алгоритми за намалување на крос-ентропија, а финалниот резултат е параметризирана сума од двете. Тренирањето и генерирањето на песните следи стандардна итеративна процедура.

@@

Јазичните модели на ниво на буква може да се прошират на повеќе начини. Во трудот \cite{Tikhonov2017} авторите предложува збогатување на самите букви и користење на embedding или врадување на самите букви и на секвенци со одредена должина. Се ограничуваат со генерирање на монофонична музика, користејќи огромна колекција на MIDI датотеки, со низа чекори за филтрирање, одстранување на сувишни информации, транспозиции и уедначувања добиле податочно множество од повеќе од 15000 песни составени само од мелодиска линија. Секоја нота од мелодиската секвенца е претставена со embedding-зи за: висината, октавата и времетраењето како и со мета-информации извлечени од MIDI датотеката. Ваквата репрезентација ја кодираат во 1 од N бинарен вектор на влез, и на излезот од системот користат 1 од N софтмакс активација. Средниот дел од архитектурата се состои од енкодер и декодер длабока рекурентна подмрежа. Двете подмрежи се обратно свртени пирамиди кои се сретнуваа со врвовите, т.е. во енкодер во секој нареден слој се намалува бројот на неврони се додека да се стигне до крајниот слој наречен тесно грлчо. Така секој нареден слој ги компресира податоците во густа репрезентација. Во декодерот се случува обратното, секој нареден слој има повеќе неврони и тој го открива значењето на густата реперзенатиција. Излезот од декодерот оди во активацискиот слој. Песните се делат на подсеквенци и моделот ги учи нивните густи репрезентации и како да ги добие секвенците од нивните репрезентации. Архитектурата ја нарекле VRASH - Variational Recurrent Autoencoder Supported By History (Варијациски аутоенкодер поддржан од историја) и претставува обид за додавање на варијациски баесов шум кон јазичен модел со користење на рекурентен аутоенкодер.

Во трудот \cite{Mehri2016} авторите предложуваат хиерархија од модели за учење на музка од аудио сигнал. Најниското ниво обработува рамки со големи од еден примерок, а секое нагорно ниво обработува се поголеми рамки, или подолги временски периоди, без преклоп. Сите нивоа освен најниското се длабоки рекурентни невронски мрежи, а најниското ниво е повеќеслојна перцептрон мрежа. Погорните слоеви влијајат на подолните, така што нивните излези се користат како влез на пониско ниво, слично на деконволуција. Најнискиот слој генерира дистрибуција врз просторот на можни примероци во наредниот временски момент. Целата хиерархија се тренира во целост крај-до-крај. Извршиле AB тест за субјектина оцена, споредувајќи го дво и трослојна варијанта на моделот со нивна квази-имплементација на WaveNet \cite{Oord2016} и едноставен рекурентен модел, од кои тврдат дека 3 слојниот модел е префериран од страна на оценувачите.

@@

\cite{Tikhonov2017} For each note we were building a note embedding that corresponded to the pitch of the note, an octave embedding that corresponded to the octave of the note and a delay embedding that corresponded to the length of the note. We were using this three embeddings and meta-information of a given MIDI track to build a concatenated note representation that was used as an input for training throughout this paper.

@@

\cite{Tikhonov2017}  1. Spliting the midi tracks and filtering only the desired ones by heuristic. 2. Removed strenght/velocity of the note being played, to focus on the melodic pattern determined by the pitches and temporal parameters of the notes and pauses in between and not performance nuances. 3. In order to make the learning state space denser we have centered the pitches throughout the dataset transposing median pitch of every track to the 4th octave. 4. We also normalized the pauses throughout the dataset in the following manner. For each track we have calculated a median pause. It is only to be expected that absolute majority of the pauses in the track were equal to the median pause multiplied with a rational coefficient (naturally 1/2 and 3/2 were especially popular in the majority of the tracks). We counted all possible pauses in every track and left only the tracks that had 11 different values of the pauses or less (the median + most popular pauses on each side of it). The tracks with higher variety of pauses were not included in the final dataset.
5. Finally to make the input diverse enough we have filtered the tracks with exceedingly small entropy.
